---
title: 'HW2: Class One Survey Analysis'
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
mainfont: Arial
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(results="hide", include=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
```
##### **INSTRUCTIONS**
##### Please complete tasks 1-7 using reproducible coding practices. Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks. 

##
##

##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-Fall-2021) (The dataset called 'Class 1 Survey Fall 2021.csv' is located in the class one folder), calling the R dataframe that you create *C1survey*.
```{r}
#read the C1survey dataset from Github
library(readr)
C1survey<-read_csv("/Users/kijohnson/Desktop/Class 1 Survey Fall 2021.csv")

```

##
##

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.    
```{r}
#a and b determine number of observations and variables in C1survey dataset
dim(C1survey)
#there are 28 observations and 27 variables
str(C1survey)
```

##
##

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed.
```{r}
#use names function to rename the 27 variables in order
names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

#create list of renamed variables in the object varlist
colnames(C1survey)
```
##
##

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset.
```{r}
sapply(C1survey, class) #solution 1 using sapply function
str(C1survey) #solution 2 using str function
table(sapply(C1survey, class)) #solution 3 is the best solution using table and sapply functions 23 character and 4 integer
summary(C1survey)
```

##
##
##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values. If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
#a. look at values for bday and bmonth
C1survey$bday
C1survey$bmonth

#b. recode twenty-eight for bday

#bday recoding method 1

#bday recoding method 1
class(C1survey$bday) #check that it recoded as character class
C1survey$bday<-ifelse(C1survey$bday=="twenty-eight", "28", C1survey$bday)
C1survey$bday #check values 
#convert to numeric for statistics
C1survey$bday<-as.numeric(C1survey$bday)
C1survey$bday #check values 

#bday recoding method 2
#class(C1survey$bday) #check that it recoded as character class
#C1survey$bday[C1survey$bday=="twenty-eight"]<-"28"
#C1survey$bday #check values 
#convert to numeric for statistics
#C1survey$bday<-as.numeric(C1survey$bday)
#C1survey$bday #check values 

#bmonth recoding
class(C1survey$bmonth)

C1survey$bmonth[C1survey$bmonth=="May"]<-"5"
C1survey$bmonth[C1survey$bmonth=="January"]<-"1"
C1survey$bmonth #check values 

#convert to numeric for statistics
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth #check values 

#c. calculate medians for bday and bmonth 
median(C1survey$bday, na.rm=TRUE) #16
median(C1survey$bmonth, na.rm=TRUE) #6.5
```
##
##
##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter=12/1 to 2/29, spring=3/1 to 5/31, summer=6/1 to 8/31, fall=9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code you turn in. 
```{r}
#a.
attach(C1survey) #use attach so dataset doesn't have to be specified over and over again
bseason[bmonth==12|bmonth==1|bmonth==2]<-"Winter" #recode months 12, 1, and 2 as winter
bseason[bmonth>2&bmonth<6]<-"Spring" #recode months 3, 4, and 5 as spring
bseason[bmonth>5&bmonth<9]<-"Summer" #recode months 6, 7, and 8 as summer
bseason[bmonth>8&bmonth<12]<-"Fall" #recode months 9, 10, and 11 as fall
bseason #check variable

#as factor
attach(C1survey) #use attach so dataset doesn't have to be specified over and over again
bseason[bmonth==12|bmonth==1|bmonth==2]<-0 #recode months 12, 1, and 2 as winter
bseason[bmonth>2&bmonth<6]<-1 #recode months 3, 4, and 5 as spring
bseason[bmonth>5&bmonth<9]<-2 #recode months 6, 7, and 8 as summer
bseason[bmonth>8&bmonth<12]<-3 #recode months 9, 10, and 11 as fall

bseason #check variable
bseason_f<-factor(bseason, labels=c("Winter", "Spring", "Summer", "Fall"))


#b. 
table(bmonth, bseason)
table(bmonth, bseason_f)

#c.
tab<-addmargins(table(bmonth, bseason), 1)
tab
tab<-addmargins(table(bmonth, bseason_f), 1)
tab
detach(C1survey)
#Births by season: Winter=7, Spring=5, Summer=6, Fall=10
```
##
##
##### 7. Pick your favorite variable to analyze, come up with a question you want to answer with that variable, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).

# no



